Interpretable Machine Learning

### 2021-06-30 chunks/story-wines.Rmd
finalizes talk

```{r load-data,include=FALSE}
library('mlr')
library('ggplot2')
library('tidyr')
library('lime')
devtools::install_github("christophM/iml", ref = "feature-effects")
library('iml')
source('../code/prepare-wine-data.R')
```

@@

@@

@@

lrns = list(lrn.ranger, lrn.lm, lrn.rpart)

tsk = makeRegrTask(data = wine, target = "quality")

bmr = benchmark(lrns, tsk, rdesc, measures = list(mae))
bmr_tab = getBMRAggrPerformances(bmr, as.df = TRUE)
#knitr::kable(bmr_tab[-1])
```

@@

set.seed(42)
sample_size = 500
wine_subset = wine[sample(1:nrow(wine), size = sample_size),]

pred = Predictor$new(mod, data = wine_subset, y = "quality")
```

## Train Random Forest to Predict Quality {.emphasizedabit}

Mean absolute error on test data (cross-validated): `r sprintf('%.2f', bmr_tab$mae.test.mean[bmr_tab$learner.id == "regr.ranger"])`

@@

```{r check-model}
preds = pred$predict(wine)
preds$actual = wine$quality
ggplot(preds, aes(x = actual, y = .prediction, group = actual)) + 
  geom_abline(intercept = 0, slope = 1, lty = 2) +
  geom_violin() +
  scale_x_continuous("Actual quality", 
    labels = 1:10, breaks = 1:10, limit = c(2.5,9.5)) + 
  scale_y_continuous("Predicted quality", labels = 1:10, breaks = 1:10, limits = c(2.5,9.5)) 
#scale_fill_gradient(low = "white", high = "darkblue") + 
```

@@

## Which features are important?

@@

## Which features are important?

```{r feature-importance}
imp = FeatureImp$new(pred, loss = "mae")
plot(imp)
```

@@

## Rule of thumb for wine quality?

@@

## Explain individual predictions

@@

>>>>

```{r load-data,include=FALSE}
library('patchwork')
library('mlr')
library('ggplot2')
library('tidyr')
library('lime')
devtools::install_github("christophM/iml", ref = "feature-effects")
library('iml')
source('../code/prepare-wine-data.R')
```

@@

## Linear Regression {.center data-background=../images/lm.png data-background-size=contain}

## Decision Tree {.center data-background=../images/tree.png data-background-size=contain}

@@

## Comparing Model Performance

Model performance was measured as mean absolute error using 5-fold cross-validation.

@@

lrns = list(lrn.ranger, lrn.lm, lrn.rpart)
test_size = 300
test_index = sample(1:nrow(wine), size = test_size)
wine_test = wine[test_index,]
wine_train = wine[setdiff(1:nrow(wine), test_index), ]

tsk = makeRegrTask(data = wine_train, target = "quality")

bmr = benchmark(lrns, tsk, rdesc, measures = list(mae))
bmr_tab = getBMRAggrPerformances(bmr, as.df = TRUE)
bmr_tab$mae.test.mean = round(bmr_tab$mae.test.mean, 2)
bmr_tab$learner.id[bmr_tab$learner.id == "regr.ranger"] =  "Random Forest"
bmr_tab$learner.id[bmr_tab$learner.id == "regr.lm"] =  "Linear Regression"
bmr_tab$learner.id[bmr_tab$learner.id == "regr.rpart"] =  "Decision Tree"
colnames(bmr_tab) = c("", "Model", "Mean Absolute Error")
knitr::kable(bmr_tab[-1])
```

@@

set.seed(42)
pred = Predictor$new(mod, data = wine_test, y = "quality")
```

@@

```{r check-model}
preds = pred$predict(wine_test)
preds$actual = wine_test$quality
ggplot(preds, aes(x = actual, y = .prediction, group = actual)) + 
  geom_abline(intercept = 0, slope = 1, lty = 2) +
  geom_violin() +
  scale_x_continuous("Actual quality", 
    labels = 1:10, breaks = 1:10, limit = c(2.5,9.5)) + 
  scale_y_continuous("Predicted quality", labels = 1:10, breaks = 1:10, limits = c(2.5,9.5)) 
#scale_fill_gradient(low = "white", high = "darkblue") + 
```

@@

# Which features are important?

@@

1. Measure model error on test data
1. Shuffle feature of interest
1. Measure model error again
1. Compute difference or ratio

```{r feature-importance}
imp = FeatureImp$new(pred, loss = "mae", compare = "difference")
plot(imp)
```

# How do features affect predictions?

ICE: Individual Conditional Expectation curves

PDP: Partial Dependence Plots

1. Choose a feature
1. Define a grid of values along that feature
1. For each value of the grid
    1. Replace value for all data points the feature value with the grid value
    1. Get the model predictions
1. Draw a line through the grid predictions -> ICE
1. Average the lines -> PDP

@@

## Interactions

```{r, out.width = "100%"}
library("patchwork")
p1 = FeatureEffect$new(pred, method = "pdp", feature = "alcohol")$plot()
p2 = FeatureEffect$new(pred, method = "pdp", feature = "volatile.acidity")$plot()
p3 = FeatureEffect$new(pred, method = "pdp", feature = c("alcohol", "volatile.acidity"))$plot()
(p1 + p2) / p3 + plot_layout(heights = c(1,3))
```

## Interactions (H-Statistic)

H-Statistic tells us how much of the effect of two is due to interactions.

```{r}
ntrct = Interaction$new(pred, feature = "alcohol")
plot(ntrct, ncols = 4)
```

# Rule of thumb for wine quality?

@@

# Explain individual predictions

@@

# How Can the Worst Wine be Improved?
